Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828

### 2016-07-14 inst/examples/manuscript-krakow.Rmd
paper from km2 to sheets

# Publishing per year & publisher	
npub <- df %>%
     	  group_by(publication_year, publisher) %>%
	  summarize(paper = sum(paper.consumption.km2, na.rm = TRUE), n = n())

@@

# Publishing per decade & publisher	
npub <- df %>%
     	  group_by(publication_decade, publisher) %>%
	  summarize(paper = sum(paper.consumption.km2, na.rm = TRUE), n = n()) %>%
	  mutate(paper.per.title = paper/n)

@@

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=10, out.width="430px"}
df2 <- df0 %>% group_by(publication_decade, catalog) %>% summarise(paper = sum(paper.consumption.km2, na.rm = TRUE), n = n())
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = catalog)) +
     geom_point(aes(shape = catalog), size = 3) +     
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle("Overall publishing activity")
print(p)
```

@@

  df <- dfs %>% group_by(publication_year) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

df <- dfl %>% group_by(publication_year, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

  df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

```{r topics2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width="430px", fig.show="hold"}
sel <- c("hautajaiset", "häät", "juhlamenot")
catalogue <- "Fennica"
df <- df.full # full time span
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
# Selected catalogue with selected topics
df <- df %>% filter(catalog == catalogue & hit)
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub)) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

  df <- dfl %>% filter(publisher == publ)
  df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(), paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

 df <- dfl %>% filter(publication_place == place)
 df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

# Selected catalogue with selected topics
df <- df %>% filter(catalog == catalogue & hit)
df <- df %>% group_by(publication_decade, publisher) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

  dfs <- df %>% filter(publication_place %in% top) %>% group_by(publication_decade, publication_place, country) %>% summarize(n = n(), paper = sum(paper.consumption.km2, na.rm = T))  %>% mutate(paper.per.document = paper/n)

@@

```{r bookmedia1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=7, fig.show="hold", out.width="430px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal)
  pagelimit = 50
  df2$document.type <- rep(NA, nrow(df2))
  df2$document.type[df2$pagecount > pagelimit] <- "book"
  df2$document.type[df2$pagecount <= pagelimit] <- "pamphlet"
  df2 <- df2[which(!is.na(df2$document.type)),]
  df2$document.type <- factor(df2$document.type)
  df2 <- df2 %>% group_by(publication_decade, document.type) %>%
                 summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  p <- ggplot(df2, aes(x = publication_decade, y = paper.consumption.km2, group = document.type))
  p <- p + geom_point(aes(shape = document.type), size = 5)
  p <- p + geom_smooth(method = "loess", color = "black")
  p <- p + ggtitle("Paper consumption per document type")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + scale_color_manual(values=c("black", "darkgray"))
  p <- p + ggtitle(catal)
  print(p)
}
```

@@

  df2 <- df2 %>% group_by(publication_decade, document.type) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n())

  if (nrow(df2) > 0) {
    p <- ggplot(df2, aes(x = publication_decade, y = paper.consumption.km2, group = document.type))
    p <- p + geom_point(aes(shape = document.type), size = 5)
    p <- p + geom_smooth(method = "loess", color = "black")
    p <- p + ggtitle("Paper consumption per document type")
    p <- p + xlab("Year")
    p <- p + ylab("Paper consumption (km2)")
    p <- p + scale_color_manual(values=c("black", "darkgray"))
    p <- p + ggtitle(paste(place, catal))
    print(p)
  }
 }
}
```

@@

```{r LIBER-11, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=8, fig.width=12, out.width="600px"}
theme_set(theme_bw(25))
library(sorvi)
dfs <- df0 %>% group_by(publication_decade) %>%
         summarize(paper.per.doc = mean(paper.consumption.km2, na.rm = T))
p <- ggplot(dfs, aes(y = paper.per.doc, x = publication_decade)) +
       geom_point() + geom_smooth(method = "loess", size = 1, color = "black") +
       xlab("Publication year") + ylab("Paper per document") +
       ggtitle("Average paper consumption per document")
print(p)
```

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="430px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  p <- ggplot(df2, aes(y = paper.consumption.km2, x = publication_decade, shape = gatherings, linetype = gatherings))
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1, color = "black")
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5), shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", catal, ")", sep = ""))
  print(p)
}
```

@@

```{r LIBER-13B, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="430px"}
for (place in sel.places) {
for (catal in unique(df0$catalog)) {
  df2 <- filter(df0, catalog == catal & publication_place == place) %>%
           group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n())
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))

  if (nrow(df2) > 0) {
  p <- ggplot(df2, aes(y = paper.consumption.km2, x = publication_decade, shape = gatherings, linetype = gatherings))
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1, color = "black")
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5), shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", place, catal, ")", sep = ""))
  print(p)
}
}
}
```

@@

```{r 201606krakow-turkuvsotherB, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=10, fig.show="hold", out.width="430px"}
df <- df0 %>%
      filter(catalog == "Fennica") %>% 
      filter(!is.na(country)) %>%
      filter(!is.na(publication_place) & !is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01)

@@

# Keep entries with known publication place and paper consumption
# Group by year/decade and Turku/Other
dfs <- df %>% group_by(publication_decade, myplace) %>%
  summarise(titles = n(), paper = sum(paper.consumption.km2, na.rm = TRUE))

p <- ggplot(dfs, aes(x = publication_decade, y = paper, fill = myplace, order = -as.numeric(myplace))) +
     geom_bar(position = "stack", stat = "identity", color = "black") +
     ggtitle("Paper consumption") +
     scale_fill_grey() +
     guides(fill = guide_legend(reverse = TRUE, title = "")) + 
     xlab("Publication decade") + ylab("Paper (km2)")
print(p)

@@

df <- df %>%	
        select(publication_decade, publisher, paper.consumption.km2)

@@

# Publishing per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% summarize(paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

# Remove NA time periods
df <- df %>% filter(!is.na(period))
df <- df %>% group_by(period) %>%
             summarise(n = n(),
	     	   paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

```{r topics-104, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("luonnontieteet","matematiikka","fysiikka","kemia","tähtitiede","maantiede","geodesia","geofysiikka","meteorologia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
  df <- df %>% filter(hit)
  df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

```{r 201606krakow-turkuvsother, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=5, fig.width=10}
df <- subset(df0, catalog == "Fennica") %>%
  filter(publication_year >= 1640 & publication_year <= 1800)%>% 
      filter(!is.na(country)) %>%
      filter(!is.na(publication_place) & !is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01)

@@

# Keep entries with known publication place and paper consumption
# Group by year/decade and Turku/Other
dfs <- df %>% group_by(publication_decade, myplace) %>%
  summarise(titles = n(), paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

df <- dfl %>% filter(catalog == catalogue)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

```{r language1-kungliga, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="150px", fig.show="hold"}
catalogue = "Kungliga"
  df <- dfl %>% filter(catalog == catalogue) %>%
     	        group_by(publication_decade, language) %>%
                summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

df <- df0 %>% filter(catalog == catalogue)%>%
  filter(publication_year >= 1640 & publication_year <= 1800)
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub) > 0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
# Selected catalogue with selected topics
df <- dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

>>>>

# Publishing per year & publisher	
npub <- df %>%
     	  group_by(publication_year, publisher) %>%
	  summarize(paper = sum(paper, na.rm = TRUE), n = n())

@@

# Publishing per decade & publisher	
npub <- df %>%
     	  group_by(publication_decade, publisher) %>%
	  summarize(paper = sum(paper, na.rm = TRUE), n = n()) %>%
	  mutate(paper.per.title = paper/n)

@@

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=10, out.width="430px"}
df2 <- df0 %>% group_by(publication_decade, catalog) %>% summarise(paper = sum(paper, na.rm = TRUE), n = n())
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = catalog)) +
     geom_point(aes(shape = catalog), size = 3) +     
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle("Overall publishing activity")
print(p)
```

@@

  df <- dfs %>% group_by(publication_year) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

df <- dfl %>% group_by(publication_year, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

  df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

```{r topics2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width="430px", fig.show="hold"}
sel <- c("hautajaiset", "häät", "juhlamenot")
catalogue <- "Fennica"
df <- df.full # full time span
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
# Selected catalogue with selected topics
df <- df %>% filter(catalog == catalogue & hit)
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub)) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

  df <- dfl %>% filter(publisher == publ)
  df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(), paper = sum(paper, na.rm = TRUE))

@@

 df <- dfl %>% filter(publication_place == place)
 df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

# Selected catalogue with selected topics
df <- df %>% filter(catalog == catalogue & hit)
df <- df %>% group_by(publication_decade, publisher) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

  dfs <- df %>% filter(publication_place %in% top) %>% group_by(publication_decade, publication_place, country) %>% summarize(n = n(), paper = sum(paper, na.rm = T))  %>% mutate(paper.per.document = paper/n)

@@

```{r bookmedia1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=7, fig.show="hold", out.width="430px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal)
  pagelimit = 50
  df2$document.type <- rep(NA, nrow(df2))
  df2$document.type[df2$pagecount > pagelimit] <- "book"
  df2$document.type[df2$pagecount <= pagelimit] <- "pamphlet"
  df2 <- df2[which(!is.na(df2$document.type)),]
  df2$document.type <- factor(df2$document.type)
  df2 <- df2 %>% group_by(publication_decade, document.type) %>%
                 summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
  p <- ggplot(df2, aes(x = publication_decade, y = paper, group = document.type))
  p <- p + geom_point(aes(shape = document.type), size = 5)
  p <- p + geom_smooth(method = "loess", color = "black")
  p <- p + ggtitle("Paper consumption per document type")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption")
  p <- p + scale_color_manual(values=c("black", "darkgray"))
  p <- p + ggtitle(catal)
  print(p)
}
```

@@

  df2 <- df2 %>% group_by(publication_decade, document.type) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n())

  if (nrow(df2) > 0) {
    p <- ggplot(df2, aes(x = publication_decade, y = paper, group = document.type))
    p <- p + geom_point(aes(shape = document.type), size = 5)
    p <- p + geom_smooth(method = "loess", color = "black")
    p <- p + ggtitle("Paper consumption per document type")
    p <- p + xlab("Year")
    p <- p + ylab("Paper consumption")
    p <- p + scale_color_manual(values=c("black", "darkgray"))
    p <- p + ggtitle(paste(place, catal))
    print(p)
  }
 }
}
```

@@

```{r LIBER-11, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=8, fig.width=12, out.width="600px"}
theme_set(theme_bw(25))
library(sorvi)
dfs <- df0 %>% group_by(publication_decade) %>%
         summarize(paper.per.doc = mean(paper, na.rm = T))
p <- ggplot(dfs, aes(y = paper.per.doc, x = publication_decade)) +
       geom_point() + geom_smooth(method = "loess", size = 1, color = "black") +
       xlab("Publication year") + ylab("Paper per document") +
       ggtitle("Average paper consumption per document")
print(p)
```

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="430px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  p <- ggplot(df2, aes(y = paper, x = publication_decade, shape = gatherings, linetype = gatherings))
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1, color = "black")
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption")
  p <- p + guides(linetype = guide_legend(keywidth = 5), shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", catal, ")", sep = ""))
  print(p)
}
```

@@

```{r LIBER-13B, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="430px"}
for (place in sel.places) {
for (catal in unique(df0$catalog)) {
  df2 <- filter(df0, catalog == catal & publication_place == place) %>%
           group_by(publication_decade, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n())
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))

  if (nrow(df2) > 0) {
  p <- ggplot(df2, aes(y = paper, x = publication_decade, shape = gatherings, linetype = gatherings))
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1, color = "black")
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption")
  p <- p + guides(linetype = guide_legend(keywidth = 5), shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", place, catal, ")", sep = ""))
  print(p)
}
}
}
```

@@

```{r 201606krakow-turkuvsotherB, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=10, fig.show="hold", out.width="430px"}
df <- df0 %>%
      filter(catalog == "Fennica") %>% 
      filter(!is.na(country)) %>%
      filter(!is.na(publication_place) & !is.na(paper) & paper > -0.01)

@@

# Keep entries with known publication place and paper consumption
# Group by year/decade and Turku/Other
dfs <- df %>% group_by(publication_decade, myplace) %>%
  summarise(titles = n(), paper = sum(paper, na.rm = TRUE))

p <- ggplot(dfs, aes(x = publication_decade, y = paper, fill = myplace, order = -as.numeric(myplace))) +
     geom_bar(position = "stack", stat = "identity", color = "black") +
     ggtitle("Paper consumption") +
     scale_fill_grey() +
     guides(fill = guide_legend(reverse = TRUE, title = "")) + 
     xlab("Publication decade") + ylab("Paper")
print(p)

@@

df <- df %>%	
        select(publication_decade, publisher, paper)

@@

# Publishing per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% summarize(paper = sum(paper, na.rm = TRUE))

@@

# Remove NA time periods
df <- df %>% filter(!is.na(period))
df <- df %>% group_by(period) %>%
             summarise(n = n(),
	     	   paper = sum(paper, na.rm = TRUE))

@@

```{r topics-104, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("luonnontieteet","matematiikka","fysiikka","kemia","tähtitiede","maantiede","geodesia","geofysiikka","meteorologia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper, na.rm = TRUE))

@@

  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
  df <- df %>% filter(hit)
  df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

```{r 201606krakow-turkuvsother, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=5, fig.width=10}
df <- subset(df0, catalog == "Fennica") %>%
  filter(publication_year >= 1640 & publication_year <= 1800)%>% 
      filter(!is.na(country)) %>%
      filter(!is.na(publication_place) & !is.na(paper) & paper > -0.01)

@@

# Keep entries with known publication place and paper consumption
# Group by year/decade and Turku/Other
dfs <- df %>% group_by(publication_decade, myplace) %>%
  summarise(titles = n(), paper = sum(paper, na.rm = TRUE))

@@

df <- dfl %>% filter(catalog == catalogue)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper, na.rm = TRUE))

@@

```{r language1-kungliga, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="150px", fig.show="hold"}
catalogue = "Kungliga"
  df <- dfl %>% filter(catalog == catalogue) %>%
     	        group_by(publication_decade, language) %>%
                summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

df <- df0 %>% filter(catalog == catalogue)%>%
  filter(publication_year >= 1640 & publication_year <= 1800)
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub) > 0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
# Selected catalogue with selected topics
df <- dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

### 2016-07-14 inst/examples/notebook.Rmd
README polish
paper from km2 to sheets

library(gridExtra)
grid.arrange(pics[[1]], pics[[2]], pics[[3]], pics[[4]], nrow = 1)
```

@@

# Remove entries outside the investigated publication periods
df <- df %>% filter(!is.na(period)) %>%
	     group_by(period) %>%
	     summarise(n = n(),
		  paper = sum(paper.consumption.km2, na.rm = TRUE)) %>%
        mutate(paper.per.title = paper/n)

@@

df <- df %>% group_by(publication_decade, publication_place) %>%
             summarise(n = n(),
       	       paper = sum(paper.consumption.km2, na.rm = TRUE),
		       publishers = length(na.omit(unique(publisher)))
		       )

@@

df <- df %>%	
  filter(publication_year >= 1640 & publication_year <= 1828)%>%
        select(publication_decade, publisher, paper.consumption.km2)

@@

  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub) > 0) {
      dflsub$language <- gsub("language.", "", lan)
      dfl <- bind_rows(dfl, dflsub)
    }
  }
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

```{r topics-103, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

```{r topics-102, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="150px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("gray", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "darkgreen"

  p <- ggplot(df2, aes(y = paper.consumption.km2,
                       x = publication_decade,
  	       shape = gatherings,
		       linetype = gatherings))		       
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1,
         aes(color = highlight, fill = highlight))
  p <- p + scale_color_manual(values = c("darkgreen", "darkgray"))
  p <- p + scale_fill_manual(values = c("darkgreen", "darkgray"))  	 
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5),
       	          shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", catal, ")", sep = ""))
  print(p)
}
```

>>>>

library(gridExtra)
grid.arrange(pics[[1]], pics[[2]], pics[[3]], pics[[4]], nrow = 1)
#grid.arrange(pics[[1]], pics[[2]], pics[[3]], nrow = 1)
```

@@

# Remove entries outside the investigated publication periods
df <- df %>% filter(!is.na(period)) %>%
	     group_by(period) %>%
	     summarise(n = n(),
		  paper = sum(paper, na.rm = TRUE)) %>%
        mutate(paper.per.title = paper/n)

@@

df <- df %>% group_by(publication_decade, publication_place) %>%
             summarise(n = n(),
       	       paper = sum(paper, na.rm = TRUE),
		       publishers = length(na.omit(unique(publisher)))
		       )

@@

df <- df %>%	
  filter(publication_year >= 1640 & publication_year <= 1828)%>%
        select(publication_decade, publisher, paper)

@@

  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub) > 0) {
      dflsub$language <- gsub("language.", "", lan)
      dfl <- bind_rows(dfl, dflsub)
    }
  }
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

```{r topics-103, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper, na.rm = TRUE))

@@

```{r topics-102, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="150px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("gray", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "darkgreen"

  p <- ggplot(df2, aes(y = paper,
                       x = publication_decade,
  	       shape = gatherings,
		       linetype = gatherings))		       
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1,
         aes(color = highlight, fill = highlight))
  p <- p + scale_color_manual(values = c("darkgreen", "darkgray"))
  p <- p + scale_fill_manual(values = c("darkgreen", "darkgray"))  	 
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption")
  p <- p + guides(linetype = guide_legend(keywidth = 5),
       	          shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", catal, ")", sep = ""))
  print(p)
}
```
